Efficiently stable presentations from error-correcting codes

### 2023-09-21 efficient_stability.tex
Typos in Section 2

In this section we give definitions associated with the notion of ``efficient stability'' used in the paper. We reformulate a few previously known results in this framework, and give some examples that will be used later on.

@@

  Here we call \emph{tracial von Neumann algebra} a pair $(\mM,\tau)$ of a von Neumann algebra $\mM$ together with a normal faithful tracial state $\tau$ on $\mM$, which we often refer to as the \emph{trace}. The main example of interest is $\mM=M_n(\C)$, the algebra $n\times n$ complex matrices, with $\tau$ the dimension-normalized trace, which we denote $\tr(M)=\frac{1}{n}\Tr(M)$. 	We write $\|x\|_\tau=\tau(x^*x)^{1/2}$ to denote the $2$-norm on $\mM$ with respect to $\tau$.

@@

Suppose given a finite presentation of a (possibly infinite) group $G$ using generators $S$ and relations $R$. Informally, we say that the presentation is \emph{stable} if any map from $S$ to e.g.\ unitaries that approximately respects the relations $R$ is close to a representation of $G$. Furthermore, we will say that the presentation is \emph{efficient} if it is stable and provides a good trade-off between its size (the number of relations used and their length) and how the closeness to a representation depends on the error in satisfying the relations $R$. All the notions referred to informally in the preceding sentences --- ``approximately,'' ``close,'' good trade-off,'' etc., can be formalized in a variety of ways, leading to often incomparable notions. Here we present the formalization that is most natural to us, and is the one suggested by applications to quantum information and complexity.

Given a set $S$, we let $\mF(S)$ denote the free group generated by $S$. We freely identify functions from $S$ to $H$, where $H$ is any group, with homomorphisms from $\mF(S)$ to $H$. If $R$ is a subset of $\mF(S)$ then the quotient of $\mF(S)$ by the normal subgroup generated by $R$ is denoted $\langle S:R\rangle$.

We start with the notion of an almost-homomorphism, which formalizes what it means for a map defined on $S$ to approximately satisfy the relations $R$. The notion we give is a small variant on the notion of $\eps$-\emph{almost homomorphism} from a finitely presented group to a unital tracial $C^*$-algebra $\mA$ introduced in~\cite[Section 2]{hadwin2018stability}. %Informally, an $\eps$-almost homomorphism of $G=\langle S:R\rangle$ is a map from $S$ to $\mU(\mA)$ that approximately respects all relations in $R$. 
We give a variant of their definition that quantifies the error in an average sense. Below, when $\mu$ is a distribution over a finite set $\mX$ and $f:\mX\to \mR$ we write $\Es{r\sim \mu} f(r)$ for the expectation of $f$ under $\mu$.

@@

We note that this notion depends on the presentation of $G$, not only on the group itself. 
When the distribution $\mu$ is uniform over the set $R$, we simply write $\eps$-homomorphism. The definition is consistent with the usual notion of a (unitary) homomorphism, which is recovered when $\eps=0$.

@@

\begin{definition}[Efficient stability]\label{def:eff-stab}
Let $G = \langle S:R\rangle $ be a finitely presented group. Let $\mC$ be a class of tracial von Neumann algebras. Let $\mu_S$ be a distribution on $S$ and $\mu_R$ a distribution on $R$. For $\delta:[0,1]\to[0,1]$ such that $\lim_{t\to 0}\delta(t)=0$ we say that the presentation $G=\langle S:R\rangle$ is $(\delta,\mu_S,\mu_R,\mC)$-stable if for every $(\cM,\tau)$ in $\mC$, every $(\eps,\mu_R)$-almost homomorphism of $G$ is $(\delta(\eps),\mu_S)$-close to a unitary representation of $G$ on some $(\mN,\tau^\mN)\in \mC$. We refer to the function $\delta$ as the ``modulus of stability'' of the presentation.
\end{definition}

@@

Many results about various notions of stability are known. \tnote{insert more discussion here, or in the introduction} We give two results that are most relevant to our focus. First of all, 
for a finite group $G$ we can always write $G=\langle S:R\rangle$ where $S = G$ and $R=\{ g\cdot h \cdot (gh)^{-1} =e \}$. We refer to this presentation as the \emph{multiplication table presentation}\tnote{was ``exhaustive''; edited terminology}. If we let $\mu_S$ and $\mu_R$ be the uniform distribution on $S$ and on $R$ respectively then Definition~\ref{def:eff-stab} reduces to a widely used notion of \emph{flexible (Hilbert-Schmidt) stability}. In particular, for finite groups the following result is known~\cite{gowers2017inverse,de2019operator}. We adopt the formulation from~\cite[Theorem 1.4]{de2022spectral}.

@@

Next we state a result from~\cite{de2022spectral} which allows to combine stability results. The results in~\cite{de2022spectral} are rather general and apply to direct products and certain central extensions of a class of finite groups. Here, we will only use the following specialization to the case of the central extension of $\Z_2^k \times \Z_2^k$ by $\{-1,1\}$ given by $\gamma(a,b)=(-1)^{a\cdot b}$, with $a\cdot b$ the inner product modulo $2$. For a measure $\mu$ on $\Z_2^k$, define its inverse spectral gap 
\[ \kappa = \max_{a\neq 0} \frac{1}{1-\Es{b\sim\mu}(-1)^{a\cdot b}}\;.\]

@@

	Before we describe some examples, we introduce the notion of a \emph{positive operator-valued measure} (POVM), or more simply \emph{measurement}. If $\mM$ is a tracial von Neumann algebra, a measurement on $\mM$ with outcome set $\mA$ is a finite collection of positive semidefinite operators $\{P_a\}_{a\in \mA}$ such that $\sum_a P_a = \Id_\mM$. A POVM is \emph{projective} if for all $a$, $P_a$ is a projection.

@@

\begin{lemma}\label{lem:pull-back}
Let  $(\mM,\tau^\mM)$ be a tracial von Neumann algebra, $P\in\mM_\infty$ a projection of finite trace, $\mN=P\mM_\infty P$ and $\tau^\mN=\tau_\infty/\tau_\infty(P)$, and $w\in P \mM_\infty \Id_\mM$ a partial isometry. Let 
\[ \eps = \max\big\{ \tau^\mM\big(\Id_\mM - w^* w\big)\,,\;\tau^\mN\big( P- w w^*\big)\big\}\;.\] 
 Then for any projective measurement $\{T_a\}_{a \in \mA}$ on $\mN$, there is a projective measurement $\{Q_a\}_{a \in \mA}$ on $\mM$ such that 
\begin{equation}
\label{eq:pull-back} \sum_{a \in \mA} \big\| Q_a - w^* T_a w\big\|_\tau^2 \,\leq \ 56\eps\;.
\end{equation}
\end{lemma}

\begin{proof}
If $\eps\geq \frac{1}{2}$ the conclusion is trivial (for a suitably large implicit constant in the $O(\cdot)$ notation in \eqref{eq:pull-back}), so assume $\eps<\frac{1}{2}$. 
Define 
\[\tilde{Q}_a = w^* T_a w  + \frac{1}{|\mA|}\big(\Id_\mM - w^* w\big) \in \mM\;.\]
Then $\{\tilde{Q}_a\}$ is a POVM on $\mM$. Moreover, 
\begin{align*}
\sum_a \tau^\mM \big( \tilde{Q}_a^2 \big) &\geq \sum_a \tau^\mM \big( \big(w^* T_a w \big)^2 \big) \\
&= \sum_a \tau^\mM \big(  w^* T_a w w^*T_a w \big)\\
&= \sum_a \tau^\mM \big(  w^* T_a  P T_a w \big) - \sum_a \tau^\mM \big( w^* T_a  ( P - w w^*) T_a w \big)\\
&\geq 1 - \eps -  \sum_a \tau_\infty \big( w^* T_a  ( P - w w^*) T_a w \big)\\
&\geq 1 - \eps -  \tau_\infty\Big(\big( P - w w^*\big)\Big(\sum_a  T_a w w^* T_a\Big)\Big)\\ 
&\geq 1- \eps- \tau_\infty\big( P- w w^*\big)\;,
\end{align*}
where the third line uses that $T_aPT_a=T_a$, $\sum_a T_a = \Id_\mN$ and the definition of $\eps$ for the first term, and for the second the fact that for $A\in\mM$, $\tau^\mM(A)=\tau_\infty(A)$ by definition of $\tau_\infty$ and the identification of $\mM$ with a ``corner'' in $\mM_\infty$, the fourth line uses cyclicity of the trace for the second, and the last uses $\|ww^*\|,\|\sum_a T_a\|_\infty\leq 1$. By assumption, 
\begin{align*}
\tau_\infty\big( P- w w^*\big) \,\leq\, \eps\, \tau_\infty(P)\,\leq \frac{\eps}{1-\eps}\;.
\end{align*}
where the last inequality is because by definition, $\tau^N(P)=1$, thus
\[1-\eps \,\leq\, \tau^\mN(ww^*) \,=\, \frac{\tau_\infty(ww^*)}{\tau_\infty(P)}\,=\, \frac{\tau_\infty(w^*w)}{\tau_\infty(P)}  \,\leq\, \frac{1}{\tau_\infty(P)}\]
since $\tau_\infty(w^* w) = \tau^\mM(w^* w)$ and $w^*w\leq I_\mM$. Overall, 
\[ \sum_a \tau^\mM \big( \tilde{Q}_a^2 \big) \,\geq\, 1-\eps-\frac{\eps}{1-\eps}\,\geq\, 1-3\eps\;.\]
To conclude we apply~\cite[Theorem 1.2]{de2021orthogonalization} to obtain a projective measurement $\{Q_a\}$ on $\mM$ such that 
\begin{equation*}
\sum_a \big\|{Q}_a - \tilde{Q}_a \big\|^2_\tau \,=\, 27\eps\;.
\end{equation*}
Finally,
\begin{align*}
\sum_a \big\|{Q}_a - w^*{T}_a w\big\|^2_\tau &= \sum_a \Big\|{Q}_a - \tilde{Q}_a  + \frac{1}{|\mA|}\big(I_\mM - w^* w\big) \Big\|^2_2\\
&\leq  \sum_a 2\big\|{Q}_a - \tilde{Q}_a\big\|_\tau^2  + 2\frac{1}{|\mA|}\big\|I_\mM - w^* w\big\|_\tau^2 \\
&\leq 54 \eps + 2 \tau^\mM( (I_\mM - w^* w)^2 ) \\
&\leq 54 \eps + 2 \tau^\mM(I_\mM - w^* w ) \\
&\leq 56 \eps\;,
\end{align*}
where the second line is by the triangle inequality, the fourth line is due to the fact that $I_\mM - w^* w$ is positive and has operator norm at most $1$, and the last line is by $\tau^\mM(I_\mM - w^* w ) \leq \eps$.
\end{proof}

@@

\begin{corollary}\label{cor:lin-test} %\tnote{Stated this for finite dimensions. Not sure what we want in the end}
Let $d\geq 1$ and $\phi:\Z_2^k \to \mU(\mM)$ be such that 
\[ \Es{x,y\in \Z_2^k} \big\| \phi(x)\phi(y)-\phi(x+y) \big\|_{\tau}^2 \,\leq\,\eps\;.\]
Then there is a %$d'=(1+O(\eps))d$, an isometry $w:\C^d \to \C^{d'}$ and a 
projective measurement $\{P_u\}_{u\in \Z_2^k}$ on $\mM$ such that 
\[ \Es{x\in \Z_2^k} \Big\| \phi(x) -\Big(\sum_u (-1)^{u\cdot x} P_u\Big)  \Big\|_{\tau}^2 \,=\, O(\eps)\;.\]
\end{corollary}

\begin{proof}
Any $\phi$ as in the corollary statement is an $(\eps,U_R)$-almost homomorphism of $\Z_2^k$ into $(\mM,\tau)$ for the multiplication table presentation. Applying Theorem~\ref{thm:gh}, $\phi$ is $O(\eps)$-close to a homomorphism from $\Z_2^k$ to some $(\mN,\tau^\mN)$. Because $\Z_2^k$ is Abelian, such a homomorphism is given by commuting unitaries $(U_x)_{x\in\Z_2^k}$ on $\mM$. Moreover, since $\Z_2^k$ is a $2$-group, each $U_x$ satisfies $U_x^2=\Id$, hence $U_x=U_x^*$.

For every $u\in  \Z_2^k$ let $Q_u = \Es{x} (-1)^{u\cdot x} U_x$. Then each $Q_u$ is a projection such that $\sum_u Q_u=\Id$. Using Lemma~\ref{lem:pull-back}, we find a projective measurement $\{P_u\}$ on $\mM$ that satisfies the conclusions of the corollary. 
\end{proof}

The multiplication table presentation of $\Z_2^k$ is not very ``efficient'', in the sense that it has as many generators as the group size, and quadratically as many relations. There is a much more ``efficient'' presentation of $\Z_2^k$, given by

\begin{equation}\label{eq:z2-efficient}
 \Z_2^k = \langle x_1,\ldots,x_k : [x_i,x_j]=e, x_i^2=e \; \forall i\neq j \rangle\;.
\end{equation}

We call this presentation ``efficient'' because it has a number of generators and relations that is far smaller than those of the multiplication table presentation (in this case, polylogarithmic in the group size). The following lemma shows that this presentation is also stable.

\begin{lemma}[Lemma 3.8 in~\cite{slofstra2019set}]\label{lem:eff-z2}
For every $k$, there is a $\delta_k = O_k(\eps)$ such that the presentation~\eqref{eq:z2-efficient} is $\delta_k$-stable. Furthermore, the close representation can be taken on the same algebra.% \hnote{Is there a reason to do $O_k(\eps)$ instead of writing out what the dependence of $k$ is in this lemma? I feel like this is an important aspect of the result (which motivates the later results...)}\tnote{Yes, the motivation was to first hide it, and then expand in the comment (which makes it clear that some $O_k(\eps)$ will always be true for a finite group, and that we care precisely about the $k$-dependence, which at first one might not ``notice''. Some of the stuff here I did not fine-tune too much the presentation of, because I thought we'll do that once we agree with the rough outline of the results and work on an intro and possible reorganization if needed. I agree it's not optimal now.}
\end{lemma}

As one would naturally expect, the dependence of $\delta_k$ on $k$ depends on the diameter of $\Z_2^k$ for the presentation~\eqref{eq:z2-efficient}. By expressing each element of $\Z_2^k$ as a product of generators in the natural way, applying Corollary~\ref{cor:lin-test} it is possible to show that $\delta_k=O(k^2\eps)$ in Lemma~\ref{lem:eff-z2}. With more work one can get $\delta_k=O(k\eps)$, see~\cite[Theorem 3.2]{chao2017overlapping}. This is nearly tight, as shown in~\cite{chao2017overlapping}.\tnote{Insert Henry's lower bound write-up around here?} In the next section we will obtain presentations of $\Z_2^k$ that have a much better size/modulus tradeoff.

To formulate the last example we recall the definition of the Pauli matrices
	\begin{equation}\label{eq:def-pauli-1} \sigma^X = \begin{pmatrix} 0 & 1 \\ 1 & 0 \end{pmatrix}\;,\qquad \sigma^Z = \begin{pmatrix} 1 & 0 \\ 0 & -1\end{pmatrix}\;,
	\end{equation}
	and more generally for $a,b\in \F_2^k$ let 
	\begin{equation}\label{eq:def-pauli-2}
	\sigma^X(a) = \bigotimes_{i=1}^t (\sigma^X)^{a_i}\qquad\text{and}\qquad\sigma^Z(b) = \bigotimes_{i=1}^t (\sigma^Z)^{b_i}\;.
	\end{equation}
These are self-adjoint unitary operators called Pauli observables. Each observable $\sigma^X(a)$ (resp. $\sigma^Z(b)$) corresponds to the \emph{Pauli measurement} $\{ \sigma^X_a \}_{a \in \F_2^k}$ (resp. $\{ \sigma^Z_b \}_{b \in \F_2^k}$) where (in a slight abuse of notation)  
	%	We slightly abuse notation and write 
	\[ \sigma^X_a = \Es{\alpha\in\F_2^k} (-1)^{a\cdot \alpha} \sigma^X(\alpha)\qquad\text{and}\qquad\sigma^Z_b = \Es{\beta\in\F_2^k} (-1)^{b\cdot\beta} \sigma^Z(\beta).\]
	It is easy to verify that $\{\sigma^X_a\}_a$ and $\{\sigma^Z_b\}_b$ are projections summing to identity.

For an integer $k\geq 1$,  the Pauli group $\pauli_k$ is the group generated by the Pauli matrices $\sigma_X(a)$, $\sigma_Z(b)$ introduced in~\eqref{eq:def-pauli-2}. It can also be defined more abstractly as follows. Let $\gamma: \Z_2^k\times \Z_2^k \to \{-1,1\}$ be given by $\gamma(a,b)=(-1)^{a\cdot b}$. Then $\pauli_k$ is the central extension of $\Z_2^k\times \Z_2^k$ by $\{-1,1\}$ given by $\gamma$. This group 
is also known as the Heisenberg group 
\[ H_{2k+1} = \left \{ \begin{pmatrix} 1 & a & c \\ 0 & I_{k \times k} & b \\ 0 & 0 & 1 \end{pmatrix} \right\}\subseteq {\mathrm{GL}}_{k+2}(\F_2)\;.\]

We consider the following presentation for the Pauli group:
\begin{align*}
 \pauli_k \,=\, \big\langle \{J\}\cup  \{(a,b): a,b\in \Z_2^k\} &: a^2=b^2=J^2=e,\; [a,J]=[b,J]=e,\\
&\qquad  (a,b)(a',b') = J^{a\cdot b'} (a',b')(a,b) \; \quad \forall a,b\big\rangle\;.
\end{align*}
This presentation is not quite the multiplication table presentation (it has $2^{2k}+1$ generators, whereas $|\pauli_k|=2^{2k+1}$), but it is not far from it.  
Applying Theorem~\ref{thm:gh} we obtain the following consequence.%, which we again state only for the case of a subclass of approximate homomorphisms and the class of finite-dimensional matrix algebras\tnote{again, not sure what we want ultimately}.

>>>>

In this section we give definitions associated with the notion of ``efficient stability'' used in the paper. We reformulate some previously known results in this framework, and give examples that will be used later on.

@@

  Here we call \emph{tracial von Neumann algebra} a pair $(\mM,\tau)$ of a von Neumann algebra $\mM$ together with a normal faithful tracial state $\tau$ on $\mM$, which we often refer to as the \emph{trace}. The main example of interest is $\mM=M_n(\C)$, the algebra of $n\times n$ complex matrices, with $\tau$ the dimension-normalized trace, which we denote $\tr(M)=\frac{1}{n}\Tr(M)$. 	We write $\|x\|_\tau=\tau(x^*x)^{1/2}$ to denote the $2$-norm on $\mM$ with respect to $\tau$.

@@

Suppose given a finite presentation of a (possibly infinite) group $G$ using generators $S$ and relations $R$. Informally, we say that the presentation is \emph{stable} if any map from $S$ to unitaries that approximately respects the relations $R$ is close, in an appropriate sense, to a representation of $G$. Furthermore, we will say that the presentation is \emph{efficient} if it is stable and provides a good trade-off between its size (the number of relations used and their length) and how the closeness to a representation depends on the error in satisfying the relations $R$. All the notions referred to informally in the preceding sentences --- ``approximately,'' ``close,'' good trade-off,'' etc., can be formalized in a variety of ways, leading to generally incomparable definitions. Here we present the formalization that is most natural to us, and is motivated by applications to quantum information and complexity.

Given a set $S$, we let $\mF(S)$ denote the free group generated by $S$. We identify functions from $S$ to $H$, where $H$ is any group, with homomorphisms from $\mF(S)$ to $H$. If $R$ is a subset of $\mF(S)$ then the quotient of $\mF(S)$ by the normal subgroup generated by $R$ is denoted $\langle S:R\rangle$.

We start with the notion of an almost-homomorphism, which formalizes what it means for a map defined on $S$ to approximately satisfy the relations $R$. The notion we give is a small variant of the notion of $\eps$-\emph{almost homomorphism} from a finitely presented group to a unital tracial $C^*$-algebra $\mA$ introduced in~\cite[Section 2]{hadwin2018stability}. %Informally, an $\eps$-almost homomorphism of $G=\langle S:R\rangle$ is a map from $S$ to $\mU(\mA)$ that approximately respects all relations in $R$. 
We give a variant of their definition that quantifies the error in an average sense. Below, when $\mu$ is a distribution over a finite set $\mX$ and $f:\mX\to \R$ we write $\Es{x\sim \mu} f(x)$ for the expectation of $f$ under $\mu$.

@@

We note that this notion depends on the presentation $\langle S:R\rangle$ of $G$, not only on the group itself. 
When the distribution $\mu$ is uniform over the set $R$, we simply write $\eps$-homomorphism. The definition is consistent with the usual notion of a (unitary) homomorphism, which is recovered when $\eps=0$.

@@

\begin{definition}[Stability]\label{def:eff-stab}
Let $G = \langle S:R\rangle $ be a finitely presented group. Let $\mC$ be a class of tracial von Neumann algebras. Let $\mu_S$ be a distribution on $S$ and $\mu_R$ a distribution on $R$. For $\delta:[0,1]\to[0,1]$ such that $\lim_{t\to 0}\delta(t)=0$ we say that the presentation $G=\langle S:R\rangle$ is $(\delta,\mu_S,\mu_R,\mC)$-stable if for every $(\cM,\tau)$ in $\mC$, every $(\eps,\mu_R)$-almost homomorphism of $G$ is $(\delta(\eps),\mu_S)$-close to a unitary representation of $G$ on some $(\mN,\tau^\mN)\in \mC$. We refer to the optimal such function $\delta$ as the ``soundness'' of the presentation.
\end{definition}

@@

Many results about various notions of stability are known. \tnote{insert more discussion here, or in the introduction} We give two results that are most relevant to our work. First of all, 
for a finite group $G$ we can always write $G=\langle S:R\rangle$ where $S = G$ and $R=\{ g\cdot h \cdot (gh)^{-1} =e \}$. We refer to this presentation as the \emph{multiplication table presentation}\tnote{was ``exhaustive''; edited terminology}. If we let $\mu_S$ and $\mu_R$ be the uniform distribution on $S$ and on $R$ respectively then Definition~\ref{def:eff-stab} reduces to a widely used notion of \emph{flexible (Hilbert-Schmidt) stability}. In particular, for finite groups the following result is known~\cite{gowers2017inverse,de2019operator}. We adopt the formulation from~\cite[Theorem 1.4]{de2022spectral}.

@@

Next we state for later use a result from~\cite{de2022spectral} which allows to combine stability results. The results in~\cite{de2022spectral} are rather general and apply to direct products and certain central extensions of a class of finite groups. Here, we will only use the following specialization to the case of the central extension of $\Z_2^k \times \Z_2^k$ by $\{-1,1\}$ given by $\gamma(a,b)=(-1)^{a\cdot b}$, with $a\cdot b$ the inner product modulo $2$. For a measure $\mu$ on $\Z_2^k$, define its inverse spectral gap 
\[ \kappa = \max_{a\neq 0} \frac{1}{1-\Es{b\sim\mu}(-1)^{a\cdot b}}\;.\]

@@

	Before giving some examples, we introduce the notion of a \emph{positive operator-valued measure} (POVM), or more simply \emph{measurement}. This is a notion that comes from quantum mechanics and will be useful to formulate some of our statements.  
  If $\mM$ is a tracial von Neumann algebra, a measurement on $\mM$ with outcome set $\mA$ is a finite collection of positive semidefinite operators $\{P_a\}_{a\in \mA}$ such that $\sum_a P_a = \Id_\mM$. A POVM is \emph{projective} if for all $a$, $P_a$ is a projection.

@@

\begin{lemma}\label{lem:pull-back}
Let  $(\mM,\tau^\mM)$ be a tracial von Neumann algebra, $P\in\mM_\infty$ a projection of finite trace, $\mN=P\mM_\infty P$ and $\tau^\mN=\tau_\infty/\tau_\infty(P)$, and $w\in P \mM_\infty \Id_\mM$ a partial isometry. Let 
\[ \eps = \max\big\{ \tau^\mM\big(\Id_\mM - w^* w\big)\,,\;\tau^\mN\big( P- w w^*\big)\big\}\;.\] 
 Then for any projective measurement $\{T_a\}_{a \in \mA}$ on $\mN$, where $\mA$ is a finite set, there is a projective measurement $\{Q_a\}_{a \in \mA}$ on $\mM$ such that 
\begin{equation}
\label{eq:pull-back} \sum_{a \in \mA} \big\| Q_a - w^* T_a w\big\|_{\tau^\mM}^2 \,\leq \ 56\eps\;.
\end{equation}
\end{lemma}

\begin{proof}
If $\eps\geq \frac{1}{2}$ the conclusion is trivial (for a suitably large implicit constant in the $O(\cdot)$ notation in \eqref{eq:pull-back}), so assume $\eps<\frac{1}{2}$. 
Define 
\[\tilde{Q}_a = w^* T_a w  + \frac{1}{|\mA|}\big(\Id_\mM - w^* w\big) \in \mM\;.\]
Then $\{\tilde{Q}_a\}$ is a POVM on $\mM$. Moreover, 
\begin{align*}
\sum_a \tau^\mM \big( \tilde{Q}_a^2 \big) &\geq \sum_a \tau^\mM \big( \big(w^* T_a w \big)^2 \big) \\
&= \sum_a \tau^\mM \big(  w^* T_a w w^*T_a w \big)\\
&= \sum_a \tau^\mM \big(  w^* T_a  P T_a w \big) - \sum_a \tau^\mM \big( w^* T_a  ( P - w w^*) T_a w \big)\\
&\geq 1 - \eps -  \sum_a \tau_\infty \big( w^* T_a  ( P - w w^*) T_a w \big)\\
&\geq 1 - \eps -  \tau_\infty\Big(\big( P - w w^*\big)\Big(\sum_a  T_a w w^* T_a\Big)\Big)\\ 
&\geq 1- \eps- \tau_\infty\big( P- w w^*\big)\;,
\end{align*}
where the third line uses that $T_aPT_a=T_a$, $\sum_a T_a = \Id_\mN$ and the definition of $\eps$ for the first term, and for the second the fact that for $A\in\mM$, $\tau^\mM(A)=\tau_\infty(A)$ by definition of $\tau_\infty$ and the identification of $\mM$ with a ``corner'' in $\mM_\infty$, the fourth line uses cyclicity of the trace for the second, and the last uses $\|ww^*\|,\|\sum_a T_a\|_\infty\leq 1$. By assumption, 
\begin{align*}
\tau_\infty\big( P- w w^*\big) \,\leq\, \eps\, \tau_\infty(P)\,\leq \frac{\eps}{1-\eps}\;,
\end{align*}
where the last inequality is because by definition, $\tau^N(P)=1$, thus
\[1-\eps \,\leq\, \tau^\mN(ww^*) \,=\, \frac{\tau_\infty(ww^*)}{\tau_\infty(P)}\,=\, \frac{\tau_\infty(w^*w)}{\tau_\infty(P)}  \,\leq\, \frac{1}{\tau_\infty(P)}\]
since $\tau_\infty(w^* w) = \tau^\mM(w^* w)$ and $w^*w\leq I_\mM$. Overall, 
\[ \sum_a \tau^\mM \big( \tilde{Q}_a^2 \big) \,\geq\, 1-\eps-\frac{\eps}{1-\eps}\,\geq\, 1-3\eps\;.\]
To conclude we apply~\cite[Theorem 1.2]{de2021orthogonalization} to obtain a projective measurement $\{Q_a\}$ on $\mM$ such that 
\begin{equation*}
\sum_a \big\|{Q}_a - \tilde{Q}_a \big\|^2_\tau \,=\, 27\eps\;.
\end{equation*}
Finally,
\begin{align*}
\sum_a \big\|{Q}_a - w^*{T}_a w\big\|^2_\tau &= \sum_a \Big\|{Q}_a - \tilde{Q}_a  + \frac{1}{|\mA|}\big(I_\mM - w^* w\big) \Big\|^2_2\\
&\leq  \sum_a 2\big\|{Q}_a - \tilde{Q}_a\big\|_\tau^2  + 2\frac{1}{|\mA|}\big\|I_\mM - w^* w\big\|_\tau^2 \\
&\leq 54 \eps + 2 \tau^\mM( (I_\mM - w^* w)^2 ) \\
&\leq 54 \eps + 2 \tau^\mM(I_\mM - w^* w ) \\
&\leq 56 \eps\;,
\end{align*}
where the second line is by the triangle inequality, the fourth line is due to the fact that $I_\mM - w^* w$ is positive and has operator norm at most $1$, and the last line is by $\tau^\mM(I_\mM - w^* w ) \leq \eps$.
\end{proof}

@@

\begin{corollary}\label{cor:lin-test} %\tnote{Stated this for finite dimensions. Not sure what we want in the end}
Let $(\mM,\tau)$ be a tracial von Neumann algebra and $\phi:\Z_2^k \to \mU(\mM)$ such that 
\[ \Es{x,y\in \Z_2^k} \big\| \phi(x)\phi(y)-\phi(x+y) \big\|_{\tau}^2 \,\leq\,\eps\;.\]
Then there is a %$d'=(1+O(\eps))d$, an isometry $w:\C^d \to \C^{d'}$ and a 
projective measurement $\{P_u\}_{u\in \Z_2^k}$ on $\mM$ such that 
\[ \Es{x\in \Z_2^k} \Big\| \phi(x) -\Big(\sum_u (-1)^{u\cdot x} P_u\Big)  \Big\|_{\tau}^2 \,=\, O(\eps)\;.\]
\end{corollary}

\begin{proof}
Any $\phi$ as in the corollary statement is an $(\eps,U_R)$-almost homomorphism of $\Z_2^k$ into $(\mM,\tau)$ for the multiplication table presentation, where $U_R$ is the uniform distibution over all relations $x+y-(x+y)=e$. Applying Theorem~\ref{thm:gh}, $\phi$ is $O(\eps)$-close to a homomorphism from $\Z_2^k$ to some $(\mN,\tau^\mN)$. Because $\Z_2^k$ is Abelian, such a homomorphism is given by commuting unitaries $(U_x)_{x\in\Z_2^k}$ on $\mM$. Moreover, since $\Z_2^k$ is a $2$-group, each $U_x$ satisfies $U_x^2=\Id$, hence $U_x=U_x^*$.

For every $u\in  \Z_2^k$ let $Q_u = \Es{x} (-1)^{u\cdot x} U_x$. Then each $Q_u$ is a projection such that $\sum_u Q_u=\Id$, and $U_x = \sum_u (-1)^{u\cdot x} Q_u$. Furthermore, it holds that 
\begin{equation}\label{eq:lin-test-1}
   \Es{x\in \Z_2^k} \Big\| \phi(x) - w^*\Big(\sum_u (-1)^{u\cdot x} Q_u\Big)w  \Big\|_{\tau}^2 \,=\, O(\eps)\;,
\end{equation}
for some partial isometry $w\in P\mM_\infty I_\mM$ as in Definition~\ref{def:close}.
Using Lemma~\ref{lem:pull-back}, we find a projective measurement $\{P_u\}$ on $\mM$ that satisfies 
\[ \sum_u \big\| P_u - w^* Q_u w \big\|_\tau^2 \,=\, O(\eps)\;.\]
Combined with~\eqref{eq:lin-test-1} and the triangle inequality, this shows the corollary. 
\end{proof}

The multiplication table presentation of $\Z_2^k$ is not very \emph{efficient}, in the sense that it has as many generators as the group size, and quadratically as many relations. There is a much more efficient presentation of $\Z_2^k$, given by

\begin{equation}\label{eq:z2-efficient}
 \Z_2^k = \langle x_1,\ldots,x_k : [x_i,x_j]=e, x_i^2=e \; \forall i\neq j \rangle\;,
\end{equation}
where $[x_i,x_j]=x_ix_jx_i^{-1}x_j^{-1}$ is the group commutator. 
We call this presentation ``efficient'' because it has a number of generators and relations that is far smaller than those of the multiplication table presentation (in this case, polylogarithmic in the group size). The following lemma shows that this presentation is also stable; however the soundness now depends on $k$.

\begin{lemma}[Lemma 3.8 in~\cite{slofstra2019set}]\label{lem:eff-z2}
Let $\mC$ be the class of tracial von Neumann algebra. Let $\mu_R$ be the equal mixture of the uniform distribution on all words $[x_i,x_j]$ ($i\neq j$) and the uniform distribution on all words $x_i^2$. Let $\mu_S$ be the uniform distribution on $\{x_1,\ldots,x_k\}$. Then 
  for every $k$, there is a $\delta_k = O_k(\eps)$ such that the presentation~\eqref{eq:z2-efficient} is $(\delta_k,\mu_S,\mu_R,\mC)$-stable. Furthermore, the close representation can be taken on the same algebra.% \hnote{Is there a reason to do $O_k(\eps)$ instead of writing out what the dependence of $k$ is in this lemma? I feel like this is an important aspect of the result (which motivates the later results...)}\tnote{Yes, the motivation was to first hide it, and then expand in the comment (which makes it clear that some $O_k(\eps)$ will always be true for a finite group, and that we care precisely about the $k$-dependence, which at first one might not ``notice''. Some of the stuff here I did not fine-tune too much the presentation of, because I thought we'll do that once we agree with the rough outline of the results and work on an intro and possible reorganization if needed. I agree it's not optimal now.}
\end{lemma}

As one would naturally expect, the dependence of $\delta_k$ on $k$ depends on the diameter of $\Z_2^k$ for the presentation~\eqref{eq:z2-efficient}. By expressing each element of $\Z_2^k$ as a product of generators in the natural way, applying Corollary~\ref{cor:lin-test} it is possible to show that $\delta_k=O(k^2\eps)$ in Lemma~\ref{lem:eff-z2}. With more work one can get $\delta_k=O(k\eps)$, see~\cite[Theorem 3.2]{chao2017overlapping}. This is nearly tight, as shown in~\cite{chao2017overlapping}.\tnote{Insert Henry's lower bound write-up around here?} In the next section we will obtain presentations of $\Z_2^k$ that have a much better size/soundness tradeoff.

To formulate the last example we recall the definition of the Pauli matrices
	\begin{equation}\label{eq:def-pauli-1} \sigma^X = \begin{pmatrix} 0 & 1 \\ 1 & 0 \end{pmatrix}\;,\qquad \sigma^Z = \begin{pmatrix} 1 & 0 \\ 0 & -1\end{pmatrix}\;,
	\end{equation}
	and more generally for $a,b\in \Z_2^k$ let 
	\begin{equation}\label{eq:def-pauli-2}
	\sigma^X(a) = \bigotimes_{i=1}^k (\sigma^X)^{a_i}\qquad\text{and}\qquad\sigma^Z(b) = \bigotimes_{i=1}^k (\sigma^Z)^{b_i}\;.
	\end{equation}
These are self-adjoint unitary operators called Pauli observables. Each observable $\sigma^X(a)$ (resp. $\sigma^Z(b)$) corresponds to the \emph{Pauli measurement} $\{ \sigma^X_a \}_{a \in \Z_2^k}$ (resp. $\{ \sigma^Z_b \}_{b \in \Z_2^k}$) where (in a slight abuse of notation)  
	%	We slightly abuse notation and write 
	\[ \sigma^X_a = \Es{\alpha\in\F_2^k} (-1)^{a\cdot \alpha} \sigma^X(\alpha)\qquad\text{and}\qquad\sigma^Z_b = \Es{\beta\in\F_2^k} (-1)^{b\cdot\beta} \sigma^Z(\beta).\]
	It is easy to verify that $\{\sigma^X_a\}_a$ and $\{\sigma^Z_b\}_b$ are projections summing to identity.

For an integer $k\geq 1$,  the Pauli group $\pauli_k$ is the group generated by the Pauli matrices $\sigma^X(a)$, $\sigma^Z(b)$ introduced in~\eqref{eq:def-pauli-2}. It can also be defined more abstractly as follows. Let $\gamma: \Z_2^k\times \Z_2^k \to \{-1,1\}$ be given by $\gamma(a,b)=(-1)^{a\cdot b}$. Then $\pauli_k$ is the central extension of $\Z_2^k\times \Z_2^k$ by $\{-1,1\}$ given by $\gamma$. This group 
is also known as the Heisenberg group 
\[ H_{2k+1} = \left \{ \begin{pmatrix} 1 & a & c \\ 0 & I_{k \times k} & b \\ 0 & 0 & 1 \end{pmatrix} \right\}\subseteq {\mathrm{GL}}_{k+2}(\F_2)\;.\]
